%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%STARt PREAMBLE
\documentclass{letter}

%Required: You must have these
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{hyperref}
\usepackage{pdflscape}
\usepackage{array}
\usepackage{authblk}
\usepackage{amsmath}
%\usepackage[backend=bibtex]{biblatex}
%Strongly recommended
 %put your figures in one place
%\SweaveOpts{prefix.string=figures/, eps=FALSE} 
%you'll want these for pretty captioning
\usepackage[small]{caption}

\setkeys{Gin}{width=0.8\textwidth} %make the figs 50 perc textwidth
\setlength{\captionmargin}{30pt}
\setlength{\abovecaptionskip}{10pt}
\setlength{\belowcaptionskip}{10pt}
% manual for caption http://www.dd.chalmers.se/latex/Docs/PDF/caption.pdf

%Optional: I like to muck with my margins and spacing in ways that LaTeX frowns on
%Here's how to do that
\topmargin -1.5cm  
\oddsidemargin -0.04cm 
\evensidemargin -0.04cm % same as oddsidemargin but for left-hand pages
\textwidth 16.59cm
\textheight 21.94cm 
%\pagestyle{empty}  % Uncomment if don't want page numbers
\parskip 7.2pt   % sets spacing between paragraphs
%\renewcommand{\baselinestretch}{1.5} 	% Uncomment for 1.5 spacing between lines
\parindent 0pt% sets leading space for paragraphs
\usepackage{setspace}
%\doublespacing
\renewcommand{\baselinestretch}{1}
\usepackage{lineno}
 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%END PREAMBLE 

%Start of the document
\begin{document}

\textbf{Question responses to GCB:}


\textbf{Running Head: 43/45 characters}\\

The environment and evolution shape forests 


\textbf{What scientific question is addressed in this manuscript? 248 characters}\\
%emw7Oct2024 -- I think we need to be more exciting! And remember we can more here than we can in the paper (e.g., population)... here's one idea. 
%What determines budburst timing across species, sites and diverse forest communities in North America? While many models critical for future forecasts assume temperature and photoperiod explain forest leafout, we test the role of populations, species and community-level variation, including the role of evolutionary history in shaping species differences. 
%Climate change is shifting the timing of spring growth, making it critical we understand the drivers of variability in events like budburst. We tested how cues and evolutionary history relates to community-level budburst and across North America. 

% dlOct20: the box only allows 250 characters, so here is my paraphrasing of the above

Most Carbon forecasts assume temperature and light explain forest leafout, but this captures little of the observed variability. We test how cues like light and temperature alongside population-, species- and community-level differences explain leafout variation. 

\textbf{What is/are the key findings that answer this question? 246 characters}\\

%emwNov2: This is cover letter so we can say site instead of population and be careful to CARRY though what you started more (cues is not defined yet so you cannot use it). 

We find species-level differences dominate in explaining leafout variation, with some of this variation explained by species evolutionary history. Temperature and light also play a major role, but we found no effect of site.  \\

\textbf{What are the novel results, ideas, or methods presented in work? 237 characters}\\
%emw7Oct2024 -- I think the fact that our model of budburst is massively incomplete should be the answer to the PREVIOUS question (and should sound much more exciting). Also, stick with evolutionary history here and don't try to mix in phylogenetic as a term if you can. For this question you can talk a little more about the vast scale of the study and your analytical approach. 
Using the first dataset that can partition variation from environmental cues, site, and communities--- spanning 47 species---we find that species level differences may be more important than temperature and light in forecasting forest phenology. \\

\textbf{Describe how your paper fits within the scope of GCB, what biological AND global change aspects does it address. 249 characters}\\

%emw7Oct2024 -- I like the idea of pointing out here that results apply beyond forests and beyond plants, but this is ALSO your best chance to say why this matters -- carbon storage, forests etc. should be here as your opening finding. 
Our work offers novel insights into how shifts in budburst will impact future forest dynamics and carbon sequestration. This work also highlights a new analytical approach that has broad applications across diverse species assemblages and phenologies.\\

\textbf{What are three most recently published papers that are relevant to this manuscript?}\\ 
\begin{enumerate}
\item Zohner, Constantin M., Leila Mirzagholi, Susanne S. Renner, Lidong Mo, Dominic Rebindaine, Raymo Bucher, Daniel Palous, Yann Vitasse, Yongshuo H. Fu, Benjamin D. Stocker, Thomas W. Crowther. 2023. Effect of climate warming on the timing of autumn leaf senescence reverses after the summer solstice. Science. 381:1-9.\\
\item Zhang, Haicheng, Isabelle Chuine, Pierre Regnier, Philippe Ciais, Wenping Yuan. 2022. Deciphering the multiple effects of climate warming on the temporal shift of leaf unfolding. Nature Climate Change. 12: 193-199.\\
\item Rebecca A. Montgomery, Karen E. Rice, Artur Stefanski, Roy L. Rich, Peter B. Reich. 2020. Phenological responses of temperate and boreal trees to warming depend on ambient spring temperatures, leaf habit, and geographic range. PNAS. 117:10397-10405
\end{enumerate}

\textbf{Reviewers: 6}
\begin{itemize}
\item Mason Heberling : heberlingm@carnegiemnh.org
\item Meredith Zettlemoyer: meredith.zettlemoyer@umontana.edu
\item Amelia Caffara: amelia.caffarra@gmail.com
\item Jason Fridley fridley@clemson.edu
\item Rong Yu---University of Wisconsin rongyu@uwm.edu
\item Andrey Malyshev
%\item Annette menzel
\end{itemize}




%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% and for Nationally Determined Contributions, National Adaptation Plans, and other climate plans within the United Nations climate process are informed by the 'best available science.'
%Questions for the group:
%1) How detailed to get in Bayesian methods/theory in introduction?

%Text to add somewhere:
% in the past, climate science and biodiversity fields have been separated, but scientists and decision-makers  are increasingly highlighting  the role nature plays in global climate systems and the critical ecosystem services provided by nature to humans \citep{cohen2016nature,nesshover2017science,USGCRP2024}

